Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-05-27 raspi_journal.tex
Defining the sections. Use acronym file. Include cross-compiling description

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{As Easy as Pi: A Network Coding Raspberry Pi Testbed}

@@

% Current address and/or shared authorship
\firstnote{Fredrik Bajers Vej 7, Room: A3-111. Aalborg, Denmark}

@@

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{The abstract}

@@

@@

Network Coding (NC) reference \cite{ahlswede2000network}.

\section{Testbed Description}
\subsection{System Overview}
\subsection{Installation and Updating}
\subsection{Platform Support: Architectures, OS, compilers and Cross-compilation}

\section{Network Codes}
\subsection{RLNC}
\subsection{S-RLNC}
\subsection{TSNC}

\section{Topologies}
\subsection{Unicast}
\subsection{Broadcast}

\section{Analyzed Metrics}
\subsection{Encoding and Decoding Speeds}
\subsection{Latency}
\subsection{Energy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

@@

\section{Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

>>>>

% *** ACRONYM PACKAGE ***
\usepackage[nolist]{acronym}
\input{acronym} % Name of acronyms file

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{Easy as Pi: A Network Coding Raspberry Pi Testbed}

@@

% Current address and/or shared authorship
\firstnote{Fredrik Bajers Vej 7, Room A3-111. Aalborg, Denmark}

@@

% Abstract (Do not use inserted blank lines, i.e. \\)
\abstract{Due to the recurrent nature }

@@

At the beginning, there was darkness and then... bang! \ac{NC}
\cite{ahlswede2000network} appeared to save us from the evilness
of routing.

General introduction. Introduction to topic addressed in the journal.
Review of the State of the Art. Specify why our approach has benefits
and which are they. Indicate contributions.

Our work is organized as follows: Section \ref{sec:testbed} introduces
the testbed system, its setup, connectivity, files configuration and
compilation of the Kodo library. Section \ref{sec:schemes} defines the
coding schemes employed in our study. Section \ref{sec:topologies} some
configurable topologies in the tested. Later, in Section \ref{sec:metrics}
we describe the considered metrics for performance comparison of the
codes deployed in the Raspberry Pi. In Section \ref{sec:results}, we show
the measurements in the Raspberry Pi of the mentioned metrics. Final
conclusions and future work are reviewed in Section \ref{sec:conclusions}.

@@

\section{Testbed Description}
\label{sec:testbed}

We proceed to define the testbed environment by each of its conforming
parts. We later indicate a procedure to set up the testbed. In this procedure,
we summarize all the details regarding on setting the connectivity and
system files for a set of Raspberry Pis in a centralized fashion.
Afterwards, we indicate how to cross-compile the Kodo library in an
easy way. Finally, we provide further information about Kodo itself in terms
of testing, other platforms supported and source code documentation
for further references.

\subsection{System Overview}
\subsection{Installation and Updating}
\subsection{Kodo cross-compilation: From your PC to the Raspberry Pi}

Besides the previous description (\textbf{Include compiling Kodo from the
RasPi from the scratch}), the testbed administrator can compile Kodo in its
personal workstation and transfer the generated binaries directly to
a path in the Raspberry Pi. To achieve this, we get a toolchain that
contains the binaries for the \texttt{raspberry-gxx49-arm-g++} compiler
for the Raspberry Pi. Therefore, we strongly recommend any testbed
administrator to do the following procedure. In what follows, we provide
the instructions considering that the NFS server uses the \texttt{\$HOME}
directory as the working directory. However, the administrator may choose
some other working directory of its preference if desired.

\begin{enumerate}

\item Download the Raspberry Pi toolchain for 64-bit Linux from: \\
\texttt{http://buildbot.steinwurf.dk/toolchains/linux/} to your
\texttt{\$HOME} directory. \\

\item Extract the downloaded file locally in the NFS server. After
this operation, there should be a new directory for the toolchain
in the server. \\

\item Add the \texttt{bin} folder of the toolchain to the \texttt{PATH}
Linux environment variable of the server. This will help the server OS
to recognize the location of the compiler command, which will be needed
later. To do so, edit the \texttt{\$HOME/.profile} to add in a newline:
\texttt{PATH="\$PATH:\$HOME/raspberry-gxx49-arm/bin"}. Save the
\texttt{\$HOME/.profile}. \\

\item Restart the server session in order for the changes made in the
previous step take effect. To verify this, open a new terminal and type:
\texttt{raspberry-gxx49-arm-g++ --version}. A correct binary installation
should return an output similar to:

\texttt{raspberry-gxx49-arm-g++ (crosstool-NG 1.21.0) 4.9.3 20150311 (prerelease)
Copyright (C) 2014 Free Software Foundation, Inc.
This is free software; see the source for copying conditions.  There is NO
warranty; not even for MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.} \\

\item Clone the Kodo repository in the server by executing: \\
\texttt{git clone git://github.com/steinwurf/kodo.git} in \texttt{\$HOME}. \\

\item Navigate to the repository and configure \texttt{waf} by typing:
\texttt{python config.py} and select the 16th ``make specification'' file
for the Raspberry Pi which is option \texttt{[16]cxx\_raspberry\_gxx49\_arm}
presented by the file.

This command configures \texttt{waf} to use the proper compiler and its
required flags to generate the binaries for the Raspberry Pi. If the
configuration was correct, the output will indicate:
\texttt{'configure' finished successfully (X.XXXs)}, where \texttt{X.XXX}
is total time in seconds for configuring the project in the server. \\

\item Execute \texttt{python waf build}. If the build process was
successful, the generated binaries for the Raspberry Pi should be located
in \texttt{build/cxx\_raspberry\_gxx49\_arm} in the Kodo repository.
\textbf{Indicate how the binary files should look like}.

Once this procedure is made, the testbed administrator can relocate the
generated binary files to Raspberry Pi's through the network as desired
by using the \texttt{scp} command during the configuration step.

\end{enumerate}

\subsection{Kodo Builds, Platform Support and Documentation}
Link to the buildbot and the relevant webpages providing a description
of them.

\section{Coding Schemes}
\label{sec:schemes}

In this section, we present a description of the considered coding schemes
measured with our the Raspberry Pi testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations.

\subsection{Random Linear Network Coding}

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding,
all the original data packets $P_j, j \in [1,g]$, each of $B$ bits, are used
to create coded packets. In the following subsections, we describe the
basic functionalities of \ac{RLNC}. Later, we mention applications that could
potentially benefit from including \ac{RLNC} as a coding scheme.

\subsubsection{Encoding}
With \ac{RLNC}, each coded packet is a random linear
combination of the original set of packets. Hence,
a \ac{l.i.} set of $g$ coded packets, $C_i,\ i \in [1,g]$
is required in order to get the original information.
Each original packet is considered as a concatenation of elements
from a \ac{GF} of a given size $q$, which we denote
$GF(q)$. To create a coded packet, a coding coefficient
$v_{i,j}$, is chosen at random from $GF(q)$ for every
packet $P_j$ and multiplied and added
following the respective \ac{GF} arithmetics. In this way,
a coded packet is:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{i,j} \otimes P_j ,\ \forall i \in [1,g]
\end{align}

To indicate which packets were used to generate a coded packet, one
form is to append its coding coefficients. In this case, the overhead
included for $C_i,\ \forall i \in [1,g]$ by the coding coefficients
is given by:
%
\begin{align} \label{eq:coded_packet}
|v_i| &= \sum_{j=1}^{g} |v_{i,j}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
To perform decoding, we define $\textbf{C} = \left[C_1 \ldots C_g \right]^T$
and $\textbf{P} = \left[P_1 \ldots P_g \right]^T$. Then, decoding
reduces to solve the linear system $\textbf{C} = \textbf{V} \cdot \textbf{P}$
using Gaussian elimination \cite{fragouli2006network}.
Here, the coding matrix $\textbf{V}$ contains any set of $g$ linearly
independent packets $C_i$ as rows as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{1,1} & \ldots & v_{1,g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g,1} & \ldots & v_{g,g} \\
\end{array}
\right]
\end{align}
%
The decoder begins to compute and remove the contributions from each
of the pivot elements, e.g. leftmost elements in the main diagonal of
\eqref{eq:coding_matrix}, to reduce $\textbf{V}$ to reduced echelon
form. In this way, it is possible to recover the original set of packets.

\subsubsection{Recoding}
Network coding allows intermediate nodes in a network to
recombine (or recode) packets from their sources whether they are
coded or not. In general, a recoded packet should be indistinguishable
from a coded one. Thus, we define a recoded packet as
$R_i$ and its corresponding encoding vector as
$w_i$ with coding coefficients $ \left[w_{i,1} \ldots w_{i,g} \right]$,
as follows:
%
\begin{align}
\label{eq:recoded_packet}
R_i = \bigoplus_{j=1}^{g} w_{i,j} \otimes C_j  ,\ \forall i \in [1,g]
\end{align}

In \eqref{eq:recoded_packet}, $w_{i,j}$ is the coding coefficient that
multiplies $C_j$, uniformly and randomly chosen from $GF(q)$. Any
decoder that collects $R_i, i \in [1,g]$ linearly independent coded packets,
with their respective $w_i$, will be able to decode the data as mentioned
before.

\subsection{Sparse Random Linear Network Coding}

In \ac{SRLNC}, instead of considering all the packets to create a coded
packet like \ac{RLNC} does, here the encoder sets some coding coefficients to
zero when creating a new coded packet. This permits to reduce the complexity
of both the encoding and decoding operations since the encoder (or decoder)
requires less operations to process the data. Therefore, a new code parameter
to control the sparsity, the code density $d$, is included to vary the
number of non-zero packet that are combined. Here are various approaches to
set the density.

A first method is to specify the number of zeros in a given
generation size and later generate coded packets with this amount. For example:
16 zeros in a coding vector of 64 packets where the position of zeros are
modified at random for every coded packet generated. Another method is to
define an average number of zeros in the coding vector. Therefore, more or
less zeros are exactly generated for each coded packet in the process.
However, if not carefully made, using a sparse coding scheme might increase
the probability of generating \ac{l.d.} packets which incurs both overhead
from redundant information and processing delay.

\subsection{Tunable Sparse Network Coding}

\section{Topologies}
\label{sec:topologies}
\subsection{Unicast}
\subsection{Broadcast}

\section{Analyzed Metrics}
\label{sec:metrics}
\subsection{Encoding and Decoding Speeds}
\subsection{Latency}
\subsection{Energy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

@@

\section{Conclusion}
\label{sec:conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Discussion}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusions}
